\section{Introduction}
\label{sec:introduction}

As neural networks are becoming more relied on for many decision processes, there has been an increased focus into
understanding how these algorithms come to a specific output.
Should such models be used to take high-stakes decisions, as it is often the case in finance or in medicine, it is
indeed crucial for these tools to provide a justification along with their results, which could be used to discuss
or even oppose a decision.
The risk is that deep learning models, often considered as $``$black boxes$"$, could produce biased and unfair behaviours
toward certain categories of society.
Such behaviour has already been noticed in opaque algorithms such as COMPAS, used to evaluate the risk of recidivism,
which was accused of being biased toward people of color~\citep{rudin2019stop}.

As a result, many methods aiming to explain a deep learning prediction have been developed, which now constitutes a
field usually called Explainable AI (XAI).
Among these methods, some of the most common are LIME~\citep{ribeiro2016should}, SHAP\citep{lundberg2017unified}, 
or Integrated Gradients~\citep{sundararajan2017axiomatic}.\footnote{
    An insightful presentation of these methods can be found at https://christophm.github.io/interpretable-ml-book.
}
Alongside this research, there has also been a drive to create libraries unifying these different methods, enabling
their use on popular deep learning libraries, as well as integrating evaluation tools to compare these XAI methods.
To this end, several pieces of software have been proposed, including SHAP~\citep{lundberg2017unified},
InterpretML~\citep{nori2019interpretml}, OmniXAI~\citep{wenzhuo2022-omnixai} or Captum~\citep{kokhlikyan2020captum}
, among many.

However, several researchers~\citep{tonekaboni2020went, crabbe2021explaining} have noticed a lack of attention toward
a specific type of data: time series.
Temporal data is nevertheless crucial in many applications: indeed, for instance, financial and medical data commonly
consist in multivariate time series.
Models which produce predictions based on this type of data therefore need a careful consideration, as these
applications often carry high-stakes decisions.
Consequently, several feature attribution methods have recently been introduced to tackle this specific
case~\citep{choi2016retain, tonekaboni2020went, crabbe2021explaining}.
Yet, to the best of our knowledge, there seems to be a lack of a unified library to regroup and evaluate these specific
methods.

As a result, we created \texttt{\detokenize{time_interpret}} (short: \texttt{tint}), a Python library designed as an
extension of Captum~\citep{kokhlikyan2020captum}.
Although this library can be used with any PyTorch~\citep{NEURIPS2019_9015} model, it has a specific focus on time series,
providing several feature attribution methods developed for this specific type of data.
\texttt{\detokenize{time_interpret}} also provides evaluation tools, whether the true attributions are known or not, as
well as several time series datasets.
It also leverages PyTorch Lightning~\citep{Falcon_PyTorch_Lightning_2019} to simplify the use of the original PyTorch
library.
As such, it provides several common PyTorch models used to handle temporal data, as well as a specific PyTorch Lightning
wrapper.

Moreover, despite this focus on time series, several components of \texttt{\detokenize{time_interpret}} have a slightly
different application.
It provides for instance several methods aiming to explain language models such a BERT\@.
Its evaluation tools can also be used with any feature attribution methods, and not just the ones implemented in
this library.

This paper aim to give a general introduction to \texttt{\detokenize{time_interpret}}.
Furthermore, several previously unpublished methods have been developed along with this library, which we also present
here.
We hope this study will give more clarity to the corresponding codebase, and will prove useful for further research
in this field.
We encourage the reader to also refer to the library documentation for more information, especially in case of
new significant releases.